---
title: 'Statistical Inference: Project 01'
author: "Thiago Veras"
date: "July 22, 2015"
output: html_document
---

## Exponential Distribution compared with Central Limit Theorem


#### Overview

In this project the exponential distribution was investigated and compared it with the Central Limit Theorem. 
The investigation was focused on the distribution of averages of 40 exponentials using lambda = 0.2 for all the simulations. Note that was necesserary to do a thousand simulations.

#### Setting Workspace
```{r setting_workspace}
#setwd("~/thiagoveras/datasciencecoursera/statinference-030/project01")
echo = TRUE
set.seed(1)
library(ggplot2)
```

#### Simulations
```{r}
# Setting the variables as defined in the problem
# Investigating the distribution of averages of 40 exponentials. 
number <- 40
# Setting the lambda = 0.2 for all of simulations.
lambda <- 0.2
# Defining the number os simulations.
number_of_simulations <- 1000

# Creating the data set
# The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter.
dataset <- matrix(rexp(number * number_of_simulations, lambda), 
                  number_of_simulations)

row_means <- apply(dataset, 1, mean)

# Converting the data to a data frame for ggplot
row_means_data_frame <-data.frame(row_means)
```

#### Sample Mean versus Theoretical Mean
```{r}
# Calculating Sample Mean
sample_mean <- mean(row_means)
# Calculating Theoretical Mean
theoretical_mean <- 1 / lambda

# Creating a Data Frame with the results
type = c("sample", "theoretical")
value = c(round(sample_mean, digits=3), 
          round(theoretical_mean, digits=3))
df_mean = data.frame(value, type)

# Ploting the results
qplot(type, value, data=df_mean, 
      main="Sample Mean versus Theoretical Mean",
      xlab="", ylab="Mean") +
      ylim(4.7,5.2)
```

The sample mean is ***`r round(sample_mean, digits=3) `*** and the theretical mean is ***`r round(theoretical_mean, digits=3)`*** . The difference is minimal as seen in the chart above.

#### Sample Variance versus Theoretical Variance
```{r}
# Calculating Sample Standard Deviation
sample_standard_deviation <- sd(row_means)
# Calculating Theoretical Standard Deviation
theoretical_standard_deviation <- ((1/lambda) * (1/sqrt(number)))
# Calculating Sample Variance
sample_variance <- var(row_means)
# Calculating Theoretical Variance
theoretical_variance <- theoretical_standard_deviation^2

# Creating a Data Frame with the results
type = c("sample", "theoretical")
value = c(round(sample_variance, digits=3), 
          round(theoretical_variance, digits=3))
df_variance = data.frame(value, type)

# Ploting the results
qplot(type, value, data=df_variance, 
      main="Sample Variance versus Theoretical Variance",
      xlab="", ylab="Variance") +
      ylim(0.600,0.650)
```

The sample variance is ***`r round(sample_variance, digits=3) `*** and the theretical variance is ***`r round(theoretical_variance, digits=3)`*** . The difference is minimal as seen in the chart above.

#### Distribution
```{r}
# Ploting the exponential distribution 
ggplot(row_means_data_frame, aes(x=row_means)) +
geom_histogram(binwidth = lambda, fill="blue", color="black", 
               aes(y = ..density..)) +
labs(title="Exponential Distribution", x="Mean", y="Density") +
geom_vline(xintercept=sample_mean, 
           size=1.0, color="black") +
stat_function(fun=dnorm, 
              args=list(mean=sample_mean, 
                        sd=sample_standard_deviation),
              size = 1.0, color = "red") +
geom_vline(xintercept=theoretical_mean, 
           size=1.0, color="yellow", linetype = "longdash") +
stat_function(fun=dnorm, 
              args=list(mean=theoretical_mean, 
                        sd=theoretical_standard_deviation),
              size = 1.0, color = "orange")
```

The Central Limit Theorem states that for a large sample size, the distribution of the sample means will approach a normal distribution. This is true for any sample of independent random variables from any population distribution, as long as the population has a finite standard deviation. 
Since these conditions apply in this project example here, then the Central Limit theorem applies, as seen in the chart above.