---
title: 'Statistical Inference: Project 01'
author: "Thiago Veras"
date: "July 22, 2015"
output: html_document
---

## Exponential Distribution compared with Central Limit Theorem

#### Overview

In this project the exponential distribution was investigated and compared it with the Central Limit Theorem. 
The investigation was focused on the distribution of averages of 40 exponentials. 
Note that was necesserary to do a thousand simulations.

#### Setting Workspace
```{r setting_workspace}
setwd("~/CESAR/BIGDATA/datasciencecoursera/statinference-030/project01")
echo = TRUE
set.seed(1)
```

#### Simulations
##### Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.
```{r}
# Setting the variables as defined in the problem
number_of_values <- 40
number_of_interations <- 0.2
number_of_simulations <- 1000

# Creating the data simulation
dataset <- matrix(rexp(number_of_values*number_of_simulations, number_of_interations), number_of_simulations)
head(dataset, n = 10)

row_means <- apply(dataset, 1, mean)
head(row_means, n = 10)

row_means_data_frame <-data.frame(row_means) # convert to data.frame for ggplot
```

#### Sample Mean versus Theoretical Mean
##### Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.
```{r}
theory_mean <- 1 / number_of_interations
theory_mean

actual_mean <- mean(row_means)
round(actual_mean, digits=3)
```

#### Sample Variance versus Theoretical Variance
##### Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.
```{r}
theory_standard_deviation <- ((1/number_of_interations) * (1/sqrt(number_of_values)))
round(theory_standard_deviation, digits=3)

actual_standard_deviation <- sd(row_means)
round(actual_standard_deviation, digits=3)

theory_variance <- theory_standard_deviation^2
round(theory_variance, digits=3)

actual_variance <- var(row_means)
round(actual_variance, digits=3)
```

#### Distribution
##### Via figures and text, explain how one can tell the distribution is approximately normal.
```{r}
library(ggplot2)
ggplot(row_means_data_frame,aes(x=row_means)) +
geom_histogram(binwidth = number_of_interations,fill="blue", color="black", aes(y = ..density..)) +
labs(title="Density of 40 Numbers from Exponential Distribution", x="Mean of 40 Selections", y="Density") +
geom_vline(xintercept=actual_mean,size=1.0, color="black") +
stat_function(fun=dnorm,args=list(mean=actual_mean, sd=actual_standard_deviation),color = "red", size = 1.0) +
geom_vline(xintercept=theory_mean, size=1.0, color="yellow", linetype = "longdash") +
stat_function(fun=dnorm,args=list(mean=theory_mean, sd=theory_standard_deviation),color = "orange", size = 1.0)

```