---
title: 'Statistical Inference: Project 02'
author: "Thiago Veras"
date: "July 23, 2015"
output: word_document
---

## Analyzing the ToothGrowth data

#### Overview

The ToothGrowth data is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

#### Setting Workspace
```{r setting_workspace}
setwd("~/CESAR/BIGDATA/datasciencecoursera/statinference-030/project02")
echo = TRUE
library(ggplot2)
library(datasets)
```

#### Loading Data
```{r loading_data}
data("ToothGrowth")
attach(ToothGrowth)
raw_data <- ToothGrowth
```

#### Performing some basic exploratory data analyses
```{r}
# Getting the structure of the data
str(raw_data)
```
- len: Tooth Length
- supp: Supplement = Orange Juice (OJ) or Vitamin C (VC) 
- dose: Dose (mg)

The raw data contains 60 observations of these 3 variables.

```{r}
# Plotting the raw data to a basic exploratory data analyses
ggplot(data = raw_data, 
       aes(x = len, y = dose, color = supp)) +
       ylim(0,2.5) +
       xlim(0, 35) +
       ylab("Dose") +
       xlab("Length") +
       geom_point() +
       theme(legend.title=element_blank())
```

```{r}
table(raw_data$dose, raw_data$supp)
```

#### Providing a basic summary of the data
```{r}
tooth_data <- raw_data
tooth_data$dose <- factor(tooth_data$dose)
summary(tooth_data)
```

#### Using confidence intervals and T tests to compare tooth growth by supp and dose 

##### Comparing tooth growth by supplement
```{r}
oj_and_vc_dose_0.5mg <- tooth_data[tooth_data$dose==0.5,]
oj_and_vc_dose_1.0mg <- tooth_data[tooth_data$dose==1.0,]
oj_and_vc_dose_2.0mg <- tooth_data[tooth_data$dose==2.0,]

# Testing oranje juice and vitamin c using 0.5 mg
t_test_dose_0.5mg <- t.test(len ~ supp, data=oj_and_vc_dose_0.5mg, var.equal=FALSE)
# Testing oranje juice and vitamin c using 1.0 mg
t_test_dose_1.0mg <- t.test(len ~ supp, data=oj_and_vc_dose_1.0mg, var.equal=FALSE)
# Testing oranje juice and vitamin c using 2.0 mg
t_test_dose_2.0mg <- t.test(len ~ supp, data=oj_and_vc_dose_2.0mg, var.equal=FALSE)
```
Results:  
***oranje juice and vitamin c using 0.5 mg*** = [ p-value: ***`r round(t_test_dose_0.5mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_dose_0.5mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_dose_0.5mg$conf.int[2], digits=3)`*** ]  
***oranje juice and vitamin c using 1.0 mg*** = [ p-value: ***`r round(t_test_dose_1.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_dose_1.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_dose_1.0mg$conf.int[2], digits=3)`*** ]  
***oranje juice and vitamin c using 2.0 mg*** = [ p-value: ***`r round(t_test_dose_2.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_dose_2.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_dose_2.0mg$conf.int[2], digits=3)`*** ]  

##### Comparing tooth growth by dose
```{r}
oj_dose_0.5mg_and_1.0mg <- tooth_data[((tooth_data$dose==0.5 | tooth_data$dose==1.0) & (tooth_data$supp=="OJ")),]
oj_dose_0.5mg_and_2.0mg <- tooth_data[((tooth_data$dose==0.5 | tooth_data$dose==2.0) & (tooth_data$supp=="OJ")),]
oj_dose_1.0mg_and_2.0mg <- tooth_data[((tooth_data$dose==1.0 | tooth_data$dose==2.0) & (tooth_data$supp=="OJ")),]
vc_dose_0.5mg_and_1.0mg <- tooth_data[((tooth_data$dose==0.5 | tooth_data$dose==1.0) & (tooth_data$supp=="VC")),]
vc_dose_0.5mg_and_2.0mg <- tooth_data[((tooth_data$dose==0.5 | tooth_data$dose==2.0) & (tooth_data$supp=="VC")),]
vc_dose_1.0mg_and_2.0mg <- tooth_data[((tooth_data$dose==1.0 | tooth_data$dose==2.0) & (tooth_data$supp=="VC")),]

# Test 0.5 mg and 1.0 mg using orange juice
t_test_oj_dose_0.5mg_and_1.0mg <- t.test(len ~ dose, data=oj_dose_0.5mg_and_1.0mg, var.equal=FALSE)
# Test 0.5 mg and 2.0 mg using orange juice
t_test_oj_dose_0.5mg_and_2.0mg <- t.test(len ~ dose, data=oj_dose_0.5mg_and_2.0mg, var.equal=FALSE)
# Test 1.0 mg and 2.0 mg using orange juice
t_test_oj_dose_1.0mg_and_2.0mg <- t.test(len ~ dose, data=oj_dose_1.0mg_and_2.0mg, var.equal=FALSE)

# Test 0.5 mg and 1.0 mg using vitamin c
t_test_vc_dose_0.5mg_and_1.0mg <- t.test(len ~ dose, data=vc_dose_0.5mg_and_1.0mg, var.equal=FALSE)
# Test 0.5 mg and 2.0 mg using vitamin c
t_test_vc_dose_0.5mg_and_2.0mg <- t.test(len ~ dose, data=vc_dose_0.5mg_and_2.0mg, var.equal=FALSE)
# Test 1.0 mg and 2.0 mg using vitamin c
t_test_vc_dose_1.0mg_and_2.0mg <- t.test(len ~ dose, data=vc_dose_1.0mg_and_2.0mg, var.equal=FALSE)
```
Results:  
***0.5 mg and 1.0 mg using orange juice*** = [ p-value: ***`r round(t_test_oj_dose_0.5mg_and_1.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_oj_dose_0.5mg_and_1.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_oj_dose_0.5mg_and_1.0mg$conf.int[2], digits=3)`*** ]   
***0.5 mg and 2.0 mg using orange juice*** = [ p-value: ***`r round(t_test_oj_dose_0.5mg_and_2.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_oj_dose_0.5mg_and_2.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_oj_dose_0.5mg_and_2.0mg$conf.int[2], digits=3)`*** ]  
***1.5 mg and 2.0 mg using orange juice*** = [ p-value: ***`r round(t_test_oj_dose_1.0mg_and_2.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_oj_dose_1.0mg_and_2.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_oj_dose_1.0mg_and_2.0mg$conf.int[2], digits=3)`*** ]  
  
***0.5 mg and 1.0 mg using vitamin c*** = [ p-value: ***`r round(t_test_vc_dose_0.5mg_and_1.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_vc_dose_0.5mg_and_1.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_vc_dose_0.5mg_and_1.0mg$conf.int[2], digits=3)`*** ]   
***0.5 mg and 2.0 mg using vitamin c*** = [ p-value: ***`r round(t_test_vc_dose_0.5mg_and_2.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_vc_dose_0.5mg_and_2.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_vc_dose_0.5mg_and_2.0mg$conf.int[2], digits=3)`*** ]  
***1.5 mg and 2.0 mg using vitamin c*** = [ p-value: ***`r round(t_test_vc_dose_1.0mg_and_2.0mg$p.value, digits=3)`*** | confidence interval (-): ***`r round(t_test_vc_dose_1.0mg_and_2.0mg$conf.int[1], digits=3)`*** | confidence interval (+): ***`r round(t_test_vc_dose_1.0mg_and_2.0mg$conf.int[2], digits=3)`*** ] 

#### Conclusion 

There are clear indications that the supplement as the dosage have independent effects on the length of teeth.
